\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-23 Manuscript/manuscript.tex
smoothed results for rank-based analysis

With our small sample size, the asymptotic normal distribution may not be
reliable, so we also use 1~million random permutations to simulate the null 
distribution of the test statistic.
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

The rank-based methods described in Section~\ref{sec:rank} are used
to determine if the reaction times for athletes who had competed at other 
championships differed significantly from the 2022 World Championships. This is
a two group comparison performed three times: comparing 2019 and 2022 World
Championships, comparing 2022 and 2023 World Championships, and comparing 2022 
national championships to the 2022 World Championships.

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
   \midrule
   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
   2022 Nat. vs Inter. Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
   2022 Nat. vs Inter. Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
   2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
   2022 vs 2023 Female & $<0.001$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}
\of{Table might still be too wide.  I used <0.001 becuase that is what we
had currently been using for any values from the permutation test where the value
was < 2.2e-16 in R.  Professor Yan will know best, but are the sample sizes now
sufficiently large enough that the permutation test is not needed? If I recall
correctlty that was a correction for small sample sizes.}

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation and asymptotic rank-based tests for the six comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023
comparisons. The highly significant permutation test results
provides substantial evidence that both men and women who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior. When looked at in context
with other tests, these results show that athletes competing at the 2022 World
Track and Field Championships had faster reaction times than when they had
competed at other competitions.

@@

\jy{Owen, please prepare the following:
1) A table of the model parameter estimates (both point estimate and
standard error). This is similar to the current table, with and
without 2022 data. Let's focus on model gg3b and leave the model
comparison results in the supplement.
2) Figure: random effects at the venue level with/without 2022
data. Similar to the current figure.
3) Diagnostic plots: qq-plot for the gg3b model. QQ-plot for the
normal random effects could be placed in supplement.
4) Table: assessing the 0.1 threshold, similar to the current table.
I'll take care of the writing when you get the results.}
\of{1) Is resolved in table 2, 2) is resolved in figure 4, 3) is resolved in
figure 3. 4) is resolved in tables 3 and 4.}

\jy{Owen: an additional figure of the GG density without random
  effects (use \texttt{curve(dGG(x, mu = , sigma = , nu =  ), 0, 0.03)}) and
  with random effects (through \texttt{density()})}
\of{Please see figure 5}
\jy{This cannot be right. Black is always above blue. Both need to
  integrate to one.}

We considered many types of models from various continuous distribution families
before ultimately deciding on the Generalized Gamma model desrcribed in
Section~\ref{sec:gamlss}.  The residuals of the GG model were particularly
strong, as evidenced by Figure~\ref{fig:Residuals}.  We care especially about
the lower residuals as we are trying to estimate the probability of a low reaction
time.  While there are two outliers present in the residual plot, we accept that
this is relatively small in relation to the rest of the roughly 770 observations.

\jy{Use aspect ratio 1 for QQ plot}

\begin{figure}[tbp]
  \centering
  \includegraphics{GGResiduals}
  \caption{Residuals from the GG model}
  \label{fig:Residuals}
\end{figure}

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the gg3b model fit}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & AIC & $\mu$ & $\sigma$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
    \midrule
    Excluding 2022 & -3660 & -1.900 & -2.192 & -1.315 & 0.0434 & 0.3259 \\
    Including 2022 & -4103 & -1.910 & -2.200 & -1.178 & 0.0576 & 0.3198 \\
    \bottomrule
  \end{tabular}
\end{table}
\of{Created these with the function I wrote model-name at bottom of gamlss.R. Do
any other parametes need to be added?}

>>>>

When the sample size is small, the asymptotic normal distribution may
not be reliable, so we also use 1~million random permutations to
simulate the null distribution of the test statistic.
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

The rank-based methods described in Section~\ref{sec:rank} were used to 
compare reaction times between the 2022 World Championships and other 
competitions in which the same athletes participated. These comparisons 
were conducted separately for men and women, resulting in six total 
comparisons: reaction times from the 2022 national-level championships 
versus the 2022 World Championships for men and women, reaction times 
from the 2019 versus 2022 World Championships for men and women, and 
reaction times from the 2022 versus 2023 World Championships for men and 
women.

\begin{table}
  \centering
  \caption{Summaries of the results of comparisons between 
    reaction times from different competitions for the same athletes. 
    2022 Nat. vs Inter. compares reaction times from 2022 national-level 
    championships and the 2022 World Track and Field Championships. 2019 
    vs 2022 compares reaction times from the 2019 and 2022 World Track and 
    Field Championships. 2022 vs 2023 compares reaction times from the 
    2022 and 2023 World Track and Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
   \midrule
   2022 Nat. vs Inter. Men & $1.0 \cdot 10^{-6}$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
   2022 Nat. vs Inter. Women & $1.0 \cdot 10^{-6}$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\[1ex]
   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
   2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}

Table~\ref{tab:Clusrankresults} presents the results from both permutation 
and asymptotic rank-based tests for these six comparisons. The tests reveal 
that the national versus international comparisons yielded the lowest 
p-values, indicating stronger evidence of faster reaction times at the 
2022 World Championships compared to the 2022 national competitions. For 
both men and women, permutation test results were highly significant, 
providing substantial evidence that athletes competing at the 2022 World 
Championships had significantly faster reaction times than they did just 
a few months earlier at national-level events. The comparisons between the 
2019 and 2022 World Championships and between the 2022 and 2023 World 
Championships also produced significant results. Collectively, 
these results suggest that athletes consistently achieved faster reaction 
times at the 2022 World Championships compared to other competitions. This 
finding supports the hypothesis that the 2022 World Championships presented 
conditions conducive to faster reaction times, potentially due to systematic 
or environmental factors.

@@

We considered many types of models from various continuous distribution families
before ultimately deciding on the generalized Gamma model desrcribed in
Section~\ref{sec:gamlss}.  The residuals of the GG model were particularly
strong, as evidenced by Figure~\ref{fig:Residuals}.  We care especially about
the lower residuals as we are trying to estimate the probability of a low reaction
time.  While there are two outliers present in the residual plot, we accept that
this is relatively small in relation to the rest of the roughly 770 observations.

\begin{figure}[tbp]
  \centering
  \includegraphics{diagnosis.pdf}
  \caption{Diagnosis of the fitted generalized Gamma distribution with
    random effects in model parameters: kernerl density of 1 million
    observations drawn from the fitted model overlaid with the
    histogram of the observed reaction times (left); Q-Q plot of the
    z-score of the quantile residuals from the fitted model.}
  \label{fig:diagnosis}
\end{figure}

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the gg3b model fit}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & AIC & $\mu$ & $\sigma$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
    \midrule
    Excluding 2022 & -3660 & -1.900 & -2.192 & -1.315 & 0.0434 & 0.3259 \\
    Including 2022 & -4103 & -1.910 & -2.200 & -1.178 & 0.0576 & 0.3198 \\
    \bottomrule
  \end{tabular}
\end{table}

\of{Created these with the function I wrote model-name at bottom of gamlss.R. Do
  any other parametes need to be added?}
\jy{Remove AIC because there is no comparison. Need their standard errors.}
